---
title: "STA 304 Final Project"
output: pdf_document
author: Mengyu Lei
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Loading Data and Required Library
The data used is collected by the US National Center for Health Statistics (NCHS).
```{r cars,echo = TRUE,message = FALSE,warning = FALSE}
rm(list = ls())
library(NHANES)
library(tidyverse)
library(sampling)
library(ggplot2)
library(gridExtra)
library(arsenal)
data("NHANESraw")
```


## Data Cleaning 
```{r clean}
### Only focus on the following variables
NHANES                    <- NHANESraw %>% filter(SurveyYr=="2011_12" & Age > 17)
NHANES                    <- na.omit(NHANES[,c(1,3,4,8:11,13,24,25,61,77)])
```

## Demographic table for the data
```{r demographic}
demographic               <- as.data.frame(summary(tableby(Smoke100 ~ ., data = NHANES[,-1])))
write.table(demographic,file = "Results/Full.Demo.csv",row.names = FALSE)
```

## Check the association between Smoke100 and BPSysAve
```{r}
P1                      <- ggplot(NHANES, aes(x = BPSysAve)) + 
                           geom_histogram(aes(y = stat(density)),binwidth = 5,fill = "#56B4E9") + 
                           geom_density(col = "red",size = 1) + 
                           theme(axis.title = element_text(size = 15), 
                                 axis.text = element_text(size = 12), 
                                 plot.title = element_text(size = 15, 
                                                           hjust = 0.5)) +
                           labs(title = "BPSysAve", y = "Density")

Count                   <- NHANES %>% group_by(Smoke100) %>% summarize(Count = n())
P2                      <- ggplot(Count, aes(x = Smoke100,y = Count)) + 
                           geom_bar(stat = "Identity",fill = "#D55E00") + 
                           theme(axis.title = element_text(size = 15), 
                                 axis.text = element_text(size = 12), 
                                 plot.title = element_text(size = 15, hjust = 0.5)) +
                           labs(title = "Smoke100")

P3                     <- ggplot(NHANES,aes(x = Smoke100, y = BPSysAve,
                                                  fill = Smoke100)) + geom_boxplot() +
                          theme(axis.title = element_text(size = 15), 
                                axis.text = element_text(size = 12), 
                                plot.title = element_text(size = 15, hjust = 0.5)) +
                          labs(title = "BPSysAve by Smoke100", y = "BPSysAve")

grid.arrange(P1,P2,P3,nrow = 1)
summary(NHANES$BPSysAve)
table(NHANES$Smoke100)
NHANES %>% group_by(Smoke100) %>% summarize(Mean = mean(BPSysAve),Median = median(BPSysAve))

### Statistics Testing
var.test(BPSysAve ~ Smoke100, data = NHANES, alternative = "two.sided")
t.test(BPSysAve ~ Smoke100, data = NHANES,var.equal = TRUE)
summary(lm(BPSysAve ~ Smoke100, data = NHANES))
```

## Predicting BPSysAve by fitting a regression model with the exposure of Smoke100
```{r model}
Lower.Model             <- lm(BPSysAve ~ Smoke100, data = NHANES)
Full.Model              <- lm(formula = as.formula(paste("BPSysAve ~",
                                                         paste(colnames(NHANES)[-c(1,9,12)],
                                                               collapse = "+"))),
                              data = NHANES)

Final.Model             <- step(Full.Model,scope = list(upper=Full.Model,lower=Lower.Model),
                                direction = "both")
summary(Final.Model)
```
## Model diagnostics
```{r diagnostic}
par(mfrow = c(2,2))
plot(Final.Model)
```

## Stratified random sampling and the stratified demographic table 
```{r strata}
ME                      <- 4
alpha                   <- 0.01
D                       <- (ME/qnorm(1 - alpha/2))^2

strata.data             <- NHANES %>% group_by(SDMVSTRA) %>% 
                           summarise(N = n(),SD = (max(BPSysAve) - min(BPSysAve))/4) %>% 
                           mutate(Cost = c(52,50,46,53,48,48,47,57,53,47,54,40,43,44))

n                       <- with(strata.data,sum(N*SD/sqrt(Cost))*sum(N*SD*sqrt(Cost))/sum(N^2*D + sum(N*SD^2)))
strata.data             <- strata.data %>% mutate(n_j = ceiling(n*(N*SD/sqrt(Cost))/(sum(N*SD/sqrt(Cost)))))

demographic             <- as.data.frame(summary(tableby(SDMVSTRA ~ ., data = NHANES[,-1])))
write.csv(demographic,file = "Results/Strata.Demo.csv",row.names = FALSE)
```

## Rerun the model with stratified sample
```{r strata model}
set.seed(1024)
strata.index            <- sampling::strata(NHANES,stratanames = "SDMVSTRA",
                                            size = strata.data$n_j,
                                            method = "srswor")
strata.nhanes           <- getdata(NHANES,strata.index)
strata.model            <- lm(BPSysAve ~ Gender + Age + Education + MaritalStatus + 
                                BPDiaAve + Smoke100  + HHIncomeMid, data = strata.nhanes)
summary(strata.model)
```

## Fit a new model for the stratified data
```{r fit}
Lower.strata            <- lm(BPSysAve ~ Smoke100, data = strata.nhanes)
Full.strata             <- lm(formula = as.formula(paste("BPSysAve ~",
                                                         paste(colnames(NHANES)[-c(1,9,12)],
                                                               collapse = "+"))),
                              data = strata.nhanes)

strata.new              <- step(Full.strata,scope = list(upper = Full.strata,
                                                         lower = Lower.strata),
                                direction = "both")
summary(strata.new)
```

## Reduced Model Comparison 
```{r reduced}
summary(lm(BPSysAve ~ Smoke100 + Age + BPDiaAve + as.factor(SDMVSTRA) , data = NHANES))
summary(lm(BPSysAve ~ Smoke100 + Age + BPDiaAve , data = strata.nhanes))
```
